---
title: "Dimensions & Units"
subtitle: "Introduction to `{ospsuite}`"
date: 2024-01-01
author: Felix MIL
institute: |
  ![](esqlabs.svg){height=100}
format: esqlabs-revealjs
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(fig.showtext = TRUE,
                      fig.align = "center")
```

## {.center}

:::center-x
 **🎯 Goal: Harmonize units and convert them easily**
:::


## In PK-Sim and MoBi

PK-Sim uses the concept of Dimensions and Units

- A **dimension** represents the type of quantity (*e.g.* Volume, Mass, Concentration)
- A **unit** is one of several ways to measure a dimension (*e.g.* L, mL; Kg, g; mg/L, µg/mL)

The software uses a **base unit** for each dimension

- Volume in L,
- Mass is in kg,
- Concentration (molar) is in µmol/L

## {.center}

:::{.callout-note}
Even if the user can deviate from base unit, `PK-Sim` will always use the base unit internally
:::

:::{.callout-important}
The base units are often different from the units that are displayed by default in PK-Sim, these are called the **display units**
:::

:::center-x
[Full List of Dimensions and Units](https://docs.open-systems-pharmacology.org/appendix/appendix#a.1.-dimensions-and-base-units)
:::

## In `{ospsuite}`

`{ospsuite}` uses the same concept of Dimensions and Units.
The package provides a set of functions for easy conversion between units of a same dimension.

- `toUnit()` to convert from one unit to another
- `toBaseUnit()` to convert from any unit to the base unit
- `toDisplayUnit()` to convert from the base unit to the display unit

```{r}
ospsuite::toUnit(quantityOrDimension = "Volume",values = 1,sourceUnit = "L", targetUnit = "mL")

ospsuite::toBaseUnit(quantityOrDimension = "Volume",values = 1000, unit = "mL")
```







## Resources

- [PK Sim section about Sensitivity Analysis](https://docs.open-systems-pharmacology.org/shared-tools-and-example-workflows/sensitivity-analysis)
- [`{ospsuite}` vignette](https://www.open-systems-pharmacology.org/OSPSuite-R/articles/sensitivity-analysis.html)
- [Presentation repository](https://github.com/esqLABS/esqlabsR-training)
